An Embedding Framework for Consistent Polyhedral Surrogates

### 2019-05-23 neurips-19/poly-embed.tex
Iterated over body of top k example, moving details to the appendix

\newcommand{\Ind}{\mathbf{1}}

@@

%\jessie{In general, we shall denote discrete reports by $r \in \R$, and real-valued reports by $u \in \U$.}
For example, 0-1 loss is a discrete loss with $\R = \Y = \{-1,1\}$
\raft{was 1,..,n but that is not consistent with the way $y$ is used in the following}
given by $\ell(r)_y = \Ind[r \neq y]$, and has Bayes risk $\risk{\ell}(p) = 1-\max_{y\in\Y} p_y$.
Two important examples of surrogates for 0-1 loss are hinge loss $L(u)_y = \max(0,1-yu)$ and logistic loss $L(u)_y = \log(1+\exp(-yu))$ for $u\in\reals$.

@@

\raf{Give it another shot, but focus on flexing our formalism.  What does hinge elicit?  How can we see that it embeds (twice) 0-1 loss from our definitions?  Then maybe one sentence for why logistic does not.}
To illustrate the idea of embeddings, we now look in detail at two key examples.
First, consider a binary classification problem on the outcomes $\{-1, +1\}$.
Both the hinge loss $L_{hinge}(u,y) := (1 - uy)_+$ and logistic loss $L_{logistic}(u,y) := \ln\left(1 + \exp(-yu)\right)$ are convex surrogates that are minimized in expectation by predicting the mode, as is 0-1 loss $\ell_{0-1}(r,y) := \Ind[r\neq y]$.
However, we say that hinge loss embeds twice the 0-1 loss, while logistic loss does not embed 0-1 loss.

@@

We conclude with a concrete example; $k=1$.
One can see that the $\ell_{top-1}$ reduces to 0-1 loss, and we characterize the distributions where $\risk{L_2}(p) = \risk{\ell_2}(p) \neq \risk{\ell_{top-k}}(p)$.

@@

We write the surrogate 
\begin{align*}
L_2(u,y) := \left( \frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+
\end{align*}
where $x_{[i]}$ is the $i^{th}$ largest component of $x$.
By Lemma~\ref{lem:top-k-optimal-corners}, we can see that for all $p \in \simplex$, there is some $u \in \{0,1\}^n$ so that $u \in \argmin_{u' \in \reals^d} \inprod{p}{L_2(u)}$.
With this, we show that $L_2$ actually embeds a discrete loss defined on the corners of the hypercube.

When embedding, we consider the discrete loss $\ell_2: \{0,1\}^n \to \reals^Y_+$
\begin{align*}
\ell_2(r,y) &= \Ind[y \not\in r] + \min\left(\frac{|r \setminus \{y\}|}{k}, 1\right)~,~
\end{align*}
and we show that $L_2$ embeds $\ell_2$ in Lemma~\ref{lem:top-k-surrogate-embeds}.
Moreover, by Proposition~\ref{prop:embed-bayes-risks}, we know $\risk{L_2} = \risk{\ell_2}$.

Finally, taking $k = 1$, we see that $\vec{0} \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ such that $\max_y p_y \leq 1/2$, while the expected loss for $\ell_{top-k}$ is strictly less when predicting $\argmax_y p_y$.
First, we see that $\ell(\vec{0}, \bar{p}) = 1$.
Taking any basis vector $e_y$, the expected loss $\ell_2(e_y, p) = 2 (1-p_y) \geq 1$ by assumption on $p$. 
For any $r$ such that $|r| \geq 2$ \jessiet{In general, $|r| \geq k+1$}, we know that  $\ell_2(r,y) \geq 1$ for all $y \in \Y$, so the expected loss much be greater than or equal to $1$.
Therefore, we have that $0 \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ such that $\max_y p_y \leq 1/2$.

So while $L_2$ given by~\citet{lapin2018analysis} is not consistent with respect to top-$k$ loss, we do show a discrete loss $\ell_2$ so that $L_2$ is consistent with respect to $\ell_2$.

@@

\section{Top-$k$}
Throughout this section, consider the surrogate and discrete loss\jessie{fill in}

@@

\begin{lemma}\label{lem:top-k-surrogate-embeds}
$L_2:\reals^n \to \reals^\Y_+$ embeds $\ell_2:2^{[n]} \to \reals^\Y_+$.
\end{lemma}
\begin{proof}
  We can verify $\ell_2(r) = L_2(\varphi(r))$ for all $r \subseteq [n]$ and $\varphi(r)$ being its conversion from a set to binary vector, so $\varphi(\R) = \{0,1\}^n$.
  Moreover, by part (1.) and the above equality, respectively, we know that $\inf_{u \in \reals^n} \inprod{p}{L_2(u)} = \min_{u \in \{0,1\}^n} \inprod{p}{L_2(u)} = \min_{r \subseteq}\ell_2(r)$, and so we conclude that $L_2$ embeds $\ell_2$.
\end{proof}

>>>>

\newcommand{\Ind}[1]{\mathbf{1}[#1]}

@@

%\jessie{In general, we shall denote discrete reports by $r \in \R$, and real-valued reports by $u \in \U$.}
For example, 0-1 loss is a discrete loss with $\R = \Y = \{-1,1\}$
\raft{was 1,..,n but that is not consistent with the way $y$ is used in the following}
given by $\ell(r)_y = \Ind{r \neq y}$, and has Bayes risk $\risk{\ell}(p) = 1-\max_{y\in\Y} p_y$.
Two important examples of surrogates for 0-1 loss are hinge loss $L(u)_y = \max(0,1-yu)$ and logistic loss $L(u)_y = \log(1+\exp(-yu))$ for $u\in\reals$.

@@

\raf{Give it another shot, but focus on flexing our formalism.  What does hinge elicit?  How can we see that it embeds (twice) 0-1 loss from our definitions?  Then maybe one sentence for why logistic does not.}
To illustrate the idea of embeddings, we now look in detail at two key examples.
First, consider a binary classification problem on the outcomes $\{-1, +1\}$.
Both the hinge loss $L_{hinge}(u,y) := (1 - uy)_+$ and logistic loss $L_{logistic}(u,y) := \ln\left(1 + \exp(-yu)\right)$ are convex surrogates that are minimized in expectation by predicting the mode, as is 0-1 loss $\ell_{0-1}(r,y) := \Ind{r\neq y}$.
However, we say that hinge loss embeds twice the 0-1 loss, while logistic loss does not embed 0-1 loss.

@@

We conclude with a concrete example; $k=1$.
One can see that the $\ell_{top-1}$ reduces to 0-1 loss, and we characterize the distributions $p \in \simplex$ where $\risk{L_2}(p) = \risk{\ell_2}(p) \neq \risk{\ell_{top-k}}(p)$.

@@

We are given the surrogate 
\begin{align}\label{eq:L-2-surrogate}
L_2(u,y) := \left( \frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+
\end{align}
where $x_{[i]}$ is the $i^{th}$ largest component of $x$.
By Lemma~\ref{lem:top-k-optimal-corners}\jessie{in the Appendix... remove ref for camera ready version}, we can see that for all $p \in \simplex$, there is some $u \in \{0,1\}^n$ so that $u \in \argmin_{u' \in \reals^d} \inprod{p}{L_2(u)}$.
With this, we show that $L_2$ actually embeds a discrete loss $\ell_2: \{0,1\}^n \to \reals^\Y_+$ defined on the corners of the hypercube below:
\begin{align}\label{eq:ell-2}
\ell_2(r,y) &= \Ind{y \not\in r} + \min\left(\frac{|r \setminus \{y\}|}{k}, 1\right)
\end{align}
We show that $L_2$ embeds $\ell_2$ in Lemma~\ref{lem:top-k-surrogate-embeds} \jessie{in the Appendix... remove ref for camera ready version}.
Moreover, by Proposition~\ref{prop:embed-bayes-risks}, we know $\risk{L_2} = \risk{\ell_2}$.

While $L_2$ given by~\citet{lapin2018analysis} is not consistent with respect to top-$k$ loss, we do show a discrete loss $\ell_2$ so that $L_2$ is consistent with respect to $\ell_2$.

Taking $k = 1$, we see that $\vec{0} \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ with $\max_y p_y \leq 1/2$, while $\risk{\ell_{top-1}}(p) = 1 - \max_y p_y < 1$.
First, we see that $\ell(\vec{0}, \bar{p}) = 1$ for all $p \in \simplex$.
Taking any basis vector $e_y$, the expected loss $\ell_2(e_y, p) = 2 (1-p_y) \geq 1$ by assumption on $p$. 
For any $r$ such that $|r| \geq 2$ \jessiet{In general, $|r| \geq k+1$}, we know that  $\ell_2(r,y) \geq 1$ for all $y \in \Y$, so the expected loss much be greater than or equal to $1$.
Therefore, we have that $\vec{0} \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ such that $\max_y p_y \leq 1/2$.

@@

\section{Top-$k$}
Throughout this section, consider the surrogate and discrete loss $L_2(u,y)~:=~\left(\frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+$ and $\ell_2(r,y) = \Ind{y \not\in r} + \min\left(\frac{|r \setminus \{y\}|}{k}, 1\right)$ given in Equations~\ref{eq:L-2-surrogate} and~\ref{eq:ell-2}, respectively.

@@

\begin{lemma}\label{lem:top-k-surrogate-embeds}
$L_2:\reals^n \to \reals^\Y_+$ embeds $\ell_2:2^{[n]} \to \reals^\Y_+$.
\end{lemma}
\begin{proof}
  We can verify $\ell_2(r) = L_2(\varphi(r))$ for all $r \subseteq [n]$ and $\varphi(r)$ being its conversion from a set to binary vector, so $\varphi(\R) = \{0,1\}^n$.
  Moreover, by Lemma~\ref{lem:top-k-optimal-corners} and the above equality, respectively, we know that $\inf_{u \in \reals^n} \inprod{p}{L_2(u)} = \min_{u \in \{0,1\}^n} \inprod{p}{L_2(u)} = \min_{r \subseteq}\ell_2(r)$, and so we conclude that $L_2$ embeds $\ell_2$.
\end{proof}
